### 2018-07-12 sections/5.res.tex

\section{Numerical Experiments} \label{sec:res}

\subsection{The incompressible Euler Equation}
\subsubsection{2D vortex merging problem}
Consider the $2D$ of the incompressible Euler equation \eqref{eq:3.8} on the square $[0,2\pi]^2$, provided with periodic boundary conditions. To treat spatial derivatives, we use Fourier spectral methods, which are equivalent to a centered finite differences scheme using as differentiation stencil the entire domain. To capture the fine details characterizing the solution, $256\times 256$ modes have been adopted. The initial condition is given in terms of the vorticity distribution $\mathbf{\omega}$. We consider the evolution of three different vortices, and the initial structure is given by
\begin{equation}\label{eqn:initial_cond_vort}
\omega = \omega_0 + \sum_{i=1}^{3} \alpha_i e^{-\dfrac{\left(x-x_i\right)^{2}+\left(y-y_i\right)^{2}}{\beta^2}}.
\end{equation}
In \eqref{eqn:initial_cond_vort}, $(x,y)$ represent the spatial coordinates, $\left( x_i, y_i\right)$ the center of the $i$-th vortex and $\alpha_i$ its maximum amplitude. In our example, we consider three vortices positioned close to the center of the domain 
\begin{equation}
\left( x_1, y_1 \right) = \left(0.75\pi,\pi\right) , \left( x_2, y_2 \right) = \left(1.25\pi,\pi\right), \left( x_3, y_3 \right) = \left(1.25\pi,1.5\pi\right),
\end{equation}
two of them having positive spin $\left( \alpha_{1,2} = \pi \right)$ and the last one rotating in the opposite direction $\left( \alpha_3 = -0.5\pi \right)$. The coefficient $\beta$, for all the vortices, is set to $\frac{1}{\pi}$. The considered triangular arrangement of vortices, is an effective initial condition for the representation of merging processes: in practice, mergers are often caused by fast-moving dipoles running into another vortex. The numerical challenge raised by this setting is the formation of thin vorticity filaments: the forces developed during the merging process drain vorticity out of the main structures into long, narrow and spiral-shaped strips of intense vorticity \cite{filaments_vort}. Given the absence of natural dissipation characterizing the Euler equation, the correct preservation of this delicate structures is necessary for an energy preserving, and consequently stable, reduced models. To have initial conditions in terms of velocity components and pressure, we define first as stream-function $\Psi$ the solution of the problem $\nabla \Psi = \omega$, and then the initial velocity is given by the curl of $\Psi$. To solve the stream-function problem, we need that the integral of $\omega$ is null: to satisfy this requirement we set $\omega_0 = 0.038$. The pressure is recovered by solving the related Poisson pressure equation. 
We use the implicit midpoint scheme as time integrator to preserve the kinetic energy of the initial condition. Having a skew-symmetric nonlinearity is of the essence: in Figure \ref{fig:unstable_full_models}, we show that formulations that are not skew symmetric lead to unstable solutions, characterized by a sudden change in the conserved quantity. The merging phenomenon is simulated for a total of $18$s using a temporal step $\Delta t=0.004$s: the same setting is adopted for simulations of reduced systems.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% Comparison Skew symmetric, divergence and advective form %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h!]
  \centering
  \begin{tikzpicture}[scale=0.5452]
    \begin{axis}[ylabel=$E_K(t)$,
                 xlabel=$t$,
                 label style={font=\large},
                 legend pos=south west,
                 legend entries={Skew Symmetric form, Divergent form, Advective form},
                 legend style={font=\large},
                 width=0.7\linewidth, 
                 height=0.5\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=4},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis]
                 \addplot[color=red,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Incompressible_Euler/Energy_compare_formulations/energy_skew_symmetric.txt};
                 \addplot[color=blue,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Incompressible_Euler/Energy_compare_formulations/energy_divergence.txt};
                 \addplot[color=black!50!green,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Incompressible_Euler/Energy_compare_formulations/energy_convective.txt};
    \end{axis}%  
  \end{tikzpicture}
  \caption{Evolution of the quadratic invariant $K(t)$ for the solution to Euler equation considering three different formulations (skew-symmetric, divergent and convective forms).}
  \label{fig:unstable_full_models}
\end{figure}
Time is the only parameter considered in the reduction and $5000$ uniformly sampled snapshots are stored. A useful information regarding system reducibility is given by the singular value behaviour and in particular by the relative information content, reported in Figure \ref{fig:RIC}. Taking the first $35$ POD modes allows to maintain more than $99\%$ of the energy of the high fidelity solution: this suggest that, given the stability of the reduction procedure, we should expect an accurate reduction if a similar number of basis functions is used. Nevertheless, also smaller sets of basis functions are considered in our simulations to study the behaviour and stability of the reduced system in underrepresented scenarios.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%% Singular values decay merging vorteces %%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h!]
\centering
\begin{tikzpicture}[scale=0.5452]
    \begin{semilogyaxis}[ylabel = $\frac{\sigma_j}{\sum_i \sigma_i}$,
                 xlabel=Basis ,
                 label style={font=\large},
                 %legend pos=north east,
                 %legend entries={$k=24$ Div,$k=102$ Div,$k=201$ Div,$k=24$ Skew,$k=102$ Skew,$k=201$ Skew},
                 legend style={font=\large},
                 grid=both,
                 ticks=both,
                 width=0.7\linewidth, 
                 height=0.5\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis,
                 ymin=0,
                 ymax = 2,
                 samples = 100]
                 \addplot[color=black,style=solid,style=ultra thick]  table[x = x, y = s] {./data/Incompressible_Euler/Singular_values/sv.txt};
    \end{semilogyaxis}
\end{tikzpicture}
\caption{Singular values decay of the snapshot matrix related to POD algorithms for the velocity components of the solution to the incompressible Euler problem.}
\label{fig:RIC}
\end{figure}
In Figure \ref{fig:approx_error_incompressible_Euler} we show that the error with respect to the high fidelity solution can be reduced, as expected, by increasing the number of basis and an arbitrary accuracy can be achieved at the cost of dealing with a more computationally expensive problem to be solved.

In Figure \ref{fig:energy_error_incompressible_Euler} we compare the quadratic invariant related to the high fidelity solution with the solution to the reduced models. As proved in section \ref{sec:skew.2}, the energy is preserved through the entire simulation for reduced models and it converges to the correct value as we enrich the approximating solution space.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%% Approximation error and energy conservation %%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h!]
\centering
\begin{subfigure}[]{0.47\linewidth}
\begin{tikzpicture}[scale=0.58]
    \begin{semilogyaxis}[ylabel = $\left | \mathbf{v}(t)-\mathbf{v}^r(t) \right |^2$,
                 xlabel=$t$,
                 label style={font=\large},
                 legend style={font=\large},
                 grid=both,
                 ticks=both,
                 width=1.4\linewidth, 
                 height=1.0\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis,
                 samples = 100,
                 cycle list name=exotic]
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = error] {./data/Incompressible_Euler/Approximation_Error/err_5.txt};
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = error] {./data/Incompressible_Euler/Approximation_Error/err_8.txt};
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = error] {./data/Incompressible_Euler/Approximation_Error/err_11.txt};
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = error] {./data/Incompressible_Euler/Approximation_Error/err_14.txt};
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = error] {./data/Incompressible_Euler/Approximation_Error/err_23.txt};
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = error] {./data/Incompressible_Euler/Approximation_Error/err_26.txt};
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = error] {./data/Incompressible_Euler/Approximation_Error/err_29.txt};
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = error] {./data/Incompressible_Euler/Approximation_Error/err_35.txt};
    \end{semilogyaxis}
\end{tikzpicture}
\caption{}
\label{fig:approx_error_incompressible_Euler}
\end{subfigure} \hfill
\begin{subfigure}[]{0.47\linewidth}
\begin{tikzpicture}[scale=0.58]
    \begin{axis}[ylabel = $|K(t)-K_r^{r}(t)|$,
                 xlabel=$t$,
                 label style={font=\large},
                 legend pos=south west,
                 legend entries={$k=5$ , $k=8$, $k=11$, $k=14$, $k=23$, $k=26$, $k=29$, $k=35$},
                 legend style={font=\large},
                 grid=both,
                 ticks=both,
                 width=1.4\linewidth, 
                 height=1.0\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis,
                 ymax = 0.5598,
                 samples = 100,
                 cycle list name=exotic]
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = error] {./data/Incompressible_Euler/Energy_conservation/ene_full_5.txt};
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = error] {./data/Incompressible_Euler/Energy_conservation/ene_full_8.txt};
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = error] {./data/Incompressible_Euler/Energy_conservation/ene_full_11.txt};
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = error] {./data/Incompressible_Euler/Energy_conservation/ene_full_14.txt};
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = error] {./data/Incompressible_Euler/Energy_conservation/ene_full_23.txt};
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = error] {./data/Incompressible_Euler/Energy_conservation/ene_full_26.txt};
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = error] {./data/Incompressible_Euler/Energy_conservation/ene_full_29.txt};
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = error] {./data/Incompressible_Euler/Energy_conservation/ene_full_35.txt};
    \end{axis}
\end{tikzpicture}
\caption{}
\label{fig:energy_error_incompressible_Euler}
\end{subfigure}
\label{fig:energy_approx_err}
\caption{(\protect\subref{fig:approx_error_incompressible_Euler}) Evolution of $L_2(t)$ error between the high fidelity and the reduced velocity for the Euler incompressible problem, with $k$ being the number of basis considered. (\protect\subref{fig:energy_error_incompressible_Euler}) Conservation of the kinetic energy for the solutions of the reduced problem for different values of $k$.}
\end{figure}
For a qualitative analysis, in Figure \ref{fig:snap_solution_incompressible_Euler}, four snapshots at different time frames are shown for the high fidelity solution and for different reduced solutions. The dynamic of the solution, in particular the formation and development of vorticity filaments, is correctly represented even with a moderate number of basis functions. With fewer bases, even though small details are not captured by the POD models, the positions and spreadings of the vorticity masses match.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%% Snapshots of merging process %%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h!]
\includegraphics[scale=0.06]{data/Incompressible_Euler/Snapshots/red_17_2.png}\hspace{1em}
\includegraphics[scale=0.06]{data/Incompressible_Euler/Snapshots/red_35_2.png}\hspace{1em}
\includegraphics[scale=0.06]{data/Incompressible_Euler/Snapshots/Full_2.png}\\

\includegraphics[scale=0.06]{data/Incompressible_Euler/Snapshots/red_17_3.png}\hspace{1em}
\includegraphics[scale=0.06]{data/Incompressible_Euler/Snapshots/red_35_3.png}\hspace{1em}
\includegraphics[scale=0.06]{data/Incompressible_Euler/Snapshots/Full_3.png}\\

\includegraphics[scale=0.06]{data/Incompressible_Euler/Snapshots/red_17_4.png}\hspace{1em}
\includegraphics[scale=0.06]{data/Incompressible_Euler/Snapshots/red_35_4.png}\hspace{1em}
\includegraphics[scale=0.06]{data/Incompressible_Euler/Snapshots/Full_4.png}\\

\includegraphics[scale=0.06]{data/Incompressible_Euler/Snapshots/red_17_5.png}\hspace{1em}
\includegraphics[scale=0.06]{data/Incompressible_Euler/Snapshots/red_35_5.png}\hspace{1em}
\includegraphics[scale=0.06]{data/Incompressible_Euler/Snapshots/Full_5.png}

\caption{Solutions of the three-vortices merging problem at $t=\left\{ 4,8,12,18 \right\}$. From left to right we have the solution of the reduced model with $k=17$, $k=35$ and the high fidelity solution.}
\label{fig:snap_solution_incompressible_Euler}
\end{figure}

\subsection{The Compressible Euler Equation}
\subsubsection{2D Kelvin-Helmholtz instability}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%% Singular values KH problem %%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{1D Shock problem}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%% Singular values 1D problem %%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\begin{tikzpicture}[scale=0.5452]
    \begin{semilogyaxis}[ylabel = $\frac{\sigma_j}{\sum_i \sigma_i}$,
                 xlabel=Basis,
                 label style={font=\large},
                 legend style={font=\large},
                 grid=both,
                 ticks=both,
                 width=0.7\linewidth, 
                 height=0.5\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis,
                 ymin=0,
                 ymax = 2,
                 samples = 100]
                 \addplot[color=black,style=solid,style=ultra thick]  table[x = n, y = sing_values] {./data/Compressible_Euler/1D/Singular_values/sv.txt};
    \end{semilogyaxis}
\end{tikzpicture}
\caption{Singular values decay of the snapshot matrix related to POD algorithms for the 1D compressible Euler problem.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%% Mass, Total momentum & Energy %%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  % MASS
  \begin{subfigure}[]{0.48\linewidth}
  \begin{tikzpicture}[scale=0.55]
    \begin{axis}[ylabel = Mass,
                 xlabel=$t$,
                 label style={font=\large},
                 legend pos=north east,
                 legend entries={$k=102$ Conv,$k=201$ Conv,$k=102$ Div,$k=201$ Div,$k=102$ Skew,$k=201$ Skew},
                 legend style={font=\large},
                 grid=both,
                 ticks=both,
                 width=1.4\linewidth, 
                 height=1.0\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis,
                 ymin=0,
                 ymax = 2,
                 samples = 100]
                 \addplot[color=red,style=dashdotted,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/convective_data/mass_102_1.txt};
                 \addplot[color=blue,style=dashdotted,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/convective_data/mass_201_1.txt};
                 \addplot[color=red,style=dashed,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/divergence_data/mass_102_1.txt};
                 \addplot[color=blue,style=dashed,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/divergence_data/mass_201_1.txt};
                 \addplot[color=red,style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/skew_symm_data/mass_102_1.txt};
                 \addplot[color=blue,style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/skew_symm_data/mass_201_1.txt};
    \end{axis}%  
  \end{tikzpicture}
  \end{subfigure}\hfill% 
  % MASS (focus)
  \begin{subfigure}[]{0.48\linewidth}
  \begin{tikzpicture}[scale=0.55]
    \begin{axis}[xlabel=$t$,
                 label style={font=\large},
                 legend pos=south west,
                 legend entries={$k=24$ Skew,$k=102$ Skew,$k=201$ Skew},
                 legend style={font=\large},
                 grid=both,
                 ticks=both,
                 width=1.4\linewidth, 
                 height=1.0\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis,
                 ymax=0.5000001]
                 \addplot[color=black!50!green,style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/skew_symm_data/mass_24_1.txt};
                 \addplot[color=red,style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/skew_symm_data/mass_102_1.txt};
                 \addplot[color=blue,style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/skew_symm_data/mass_201_1.txt};
    \end{axis}%  
  \end{tikzpicture}
  \end{subfigure}
 % TOTAL MOMENTUM
  \begin{subfigure}[]{0.48\linewidth}
  \begin{tikzpicture}[scale=0.55]
    \begin{axis}[ylabel = Momentum,
                 xlabel=$t$,
                 label style={font=\large},
                 legend pos=north east,
                 %legend entries={$k=24$ Div,$k=102$ Div,$k=201$ Div,$k=24$ Skew,$k=102$ Skew,$k=201$ Skew},
                 legend style={font=\large},
                 grid=both,
                 ticks=both,
                 width=1.4\linewidth, 
                 height=1.0\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis,
                 ymin=1,
                 ymax = 2,
                 samples = 100]
                 \addplot[color=red,style=dashdotted,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/convective_data/momentum_102_1.txt};
                 \addplot[color=blue,style=dashdotted,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/convective_data/momentum_201_1.txt};
                 \addplot[color=red,style=dashed,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/divergence_data/momentum_102_1.txt};
                 \addplot[color=blue,style=dashed,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/divergence_data/momentum_201_1.txt};
                 \addplot[color=red,style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/skew_symm_data/momentum_102_1.txt};
                 \addplot[color=blue,style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/skew_symm_data/momentum_201_1.txt};
    \end{axis}%  
  \end{tikzpicture} 
  \end{subfigure}\hfill% 
 % TOTAL MOMENTUM (focus)
  \begin{subfigure}[]{0.48\linewidth}
  \begin{tikzpicture}[scale=0.55]
    \begin{axis}[xlabel=$t$,
                 label style={font=\large},
                 legend pos=south west,
                 legend style={font=\large},
                 grid=both,
                 ticks=both,
                 width=1.4\linewidth, 
                 height=1.0\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 ymax=1.04963,
                 scale only axis]
                 \addplot[color=black!50!green,style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/skew_symm_data/momentum_24_1.txt};
                 \addplot[color=red,style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/skew_symm_data/momentum_102_1.txt};
                 \addplot[color=blue,style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/skew_symm_data/momentum_201_1.txt};
    \end{axis}%  
  \end{tikzpicture}
  \end{subfigure}
 % ENERGY
  \begin{subfigure}[]{0.48\linewidth}
  \begin{tikzpicture}[scale=0.55]
    \begin{axis}[ylabel = Energy,
                 xlabel=$t$,
                 label style={font=\large},
                 legend pos=north east,
                 %legend entries={$k=24$ Div,$k=102$ Div,$k=201$ Div,$k=24$ Skew,$k=102$ Skew,$k=201$ Skew},
                 legend style={font=\large},
                 grid=both,
                 ticks=both,
                 width=1.4\linewidth, 
                 height=1.0\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis,
                 ymin=2.5,
                 ymax = 4.5,
                 samples = 100]
                 \addplot[color=red,style=dashed,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/convective_data/energy_102_1.txt};
                 \addplot[color=blue,style=dashed,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/convective_data/energy_201_1.txt};                
                 \addplot[color=red,style=dashed,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/divergence_data/energy_102_1.txt};
                 \addplot[color=blue,style=dashed,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/divergence_data/energy_201_1.txt};
                 \addplot[color=red,style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/skew_symm_data/energy_102_1.txt};
                 \addplot[color=blue,style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/skew_symm_data/energy_201_1.txt};
    \end{axis}%  
  \end{tikzpicture} 
  \caption{}
  \label{conservation_1D}
  \end{subfigure}\hfill% 
 % ENERGY (focus)
  \begin{subfigure}[]{0.48\linewidth}
  \begin{tikzpicture}[scale=0.55]
    \begin{axis}[xlabel=$t$,
                 label style={font=\large},
                 legend pos=south west,
                 legend style={font=\large},
                 grid=both,
                 ticks=both,
                 width=1.4\linewidth, 
                 height=1.0\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis,
                 ymax = 2.87438]
                 \addplot[color=black!50!green,style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/skew_symm_data/energy_24_1.txt};
                 \addplot[color=red,style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/skew_symm_data/energy_102_1.txt};
                 \addplot[color=blue,style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/skew_symm_data/energy_201_1.txt};
    \end{axis}%  
  \end{tikzpicture}
  \caption{}
  \label{conservation_1D_focus}
  \end{subfigure}
  \caption{(\protect\subref{conservation_1D})  Evolution of the three conserved quantities for the reduced solution of the compressible Euler equation (mass, total momentum and total energy).  The divergent, advective and skew-symmetric formulations have been considered and $k=102,204$ basis are used for the reduction. (\protect\subref{conservation_1D_focus}) Focus on the conserved quantity in case of stable reduction using the skew-symmetric formulation.} 
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%% Approximation Error %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\begin{subfigure}[]{0.48\linewidth}
  \begin{tikzpicture}[scale=0.55]
    \begin{semilogyaxis}[xlabel=$t$,
                 ylabel=$e(t)$
                 label style={font=\large},
                 legend pos=outer north east,
                 legend style={font=\large},
                 legend entries={$k=75$,$k=126$,$k=177$,$k=225$,$k=276$,$k=327$,$k=375$,$k=426$},
                 grid=both,
                 ticks=both,
                 width=1.4\linewidth, 
                 height=1.0\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis,
                 cycle list name=exotic]
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/error/error_75.txt};  
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/error/error_126.txt};      
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/error/error_177.txt}; 
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/error/error_225.txt};      
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/error/error_276.txt};  
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/error/error_327.txt}; 
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/error/error_375.txt};      
                 \addplot+[style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/error/error_426.txt};              
    \end{semilogyaxis}%  
  \end{tikzpicture}
\end{subfigure}
\caption{Evolution in time of the error  between the high fidelity solution of the 1D compressible Euler and the reduced solution for different number of basis $k$. As error measure we consider $e(t)=\sqrt{\|\mathbf{r}-\mathbf{r}^r\|^2+\|\mathbf{ur}-\mathbf{ur}^r\|^2\|\mathbf{p}-\mathbf{p}^r\|^2}$}.
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%% Snapshot of pressure and density %%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\begin{subfigure}[]{0.47\linewidth}
\begin{tikzpicture}[scale=0.58]
    \begin{axis}[xlabel=$x$,
                 ylabel=Density,
                 label style={font=\large},
                 legend pos=south west,
                 legend entries={HF,$k=24$ Div,$k=102$ Div,$k=201$ Div,$k=24$ Skew,$k=102$ Skew,$k=201$ Skew},
                 legend style={font=\large},
                 grid=both,
                 ticks=both,
                 width=1.4\linewidth, 
                 height=1.0\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis]
                 \addplot[color=black,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_exact_300.txt};
                 \addplot[color=red,style=dashed,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_approx_unstable_24_300.txt};
                 \addplot[color=blue,style=dashed,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_approx_unstable_102_300.txt};
                 \addplot[color=black!50!green,style=dashed,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_approx_unstable_201_300.txt};
                 \addplot[color=red,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_approx_24_300.txt};
                 \addplot[color=blue,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_approx_102_300.txt};
                 \addplot[color=black!50!green,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_approx_201_300.txt};
    \end{axis}% 
\end{tikzpicture}
\end{subfigure} 
\begin{subfigure}[]{0.47\linewidth}
\begin{tikzpicture}[scale=0.58]
    \begin{axis}[xlabel=$x$,
                 ylabel=Pressure,
                 label style={font=\large},
                 legend pos=south west,
                 %legend entries={HF,$k=24$ Div,$k=102$ Div,$k=201$ Div,$k=24$ Skew,$k=102$ Skew,$k=201$ Skew},
                 legend style={font=\large},
                 grid=both,
                 ticks=both,
                 width=1.4\linewidth, 
                 height=1.0\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis]
                 \addplot[color=black,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_exact_300.txt};
                 \addplot[color=red,style=dashed,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_approx_unstable_24_300.txt};
                 \addplot[color=blue,style=dashed,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_approx_unstable_102_300.txt};
                 \addplot[color=black!50!green,style=dashed,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_approx_unstable_201_300.txt};
                 \addplot[color=red,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_approx_24_300.txt};
                 \addplot[color=blue,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_approx_102_300.txt};
                 \addplot[color=black!50!green,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_approx_201_300.txt};
    \end{axis}% 
\end{tikzpicture}
\label{density_reduction}
\end{subfigure}

\begin{subfigure}[]{0.47\linewidth}
\begin{tikzpicture}[scale=0.58]
    \begin{axis}[xlabel=$x$,
                 ylabel=Density,
                 label style={font=\large},
                 legend pos=south west,
                 legend entries={HF,$k=24$ Div,$k=102$ Div,$k=201$ Div,$k=24$ Skew,$k=102$ Skew,$k=201$ Skew},
                 legend style={font=\large},
                 grid=both,
                 ticks=both,
                 width=1.4\linewidth, 
                 height=1.0\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis]
                 \addplot[color=black,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_exact_1000.txt};
                 \addplot[color=red,style=dashed,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_approx_unstable_24_1000.txt};
                 \addplot[color=blue,style=dashed,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_approx_unstable_102_1000.txt};
                 \addplot[color=black!50!green,style=dashed,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_approx_unstable_201_1000.txt};
                 \addplot[color=red,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_approx_24_1000.txt};
                 \addplot[color=blue,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_approx_102_1000.txt};
                 \addplot[color=black!50!green,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_approx_201_1000.txt};
    \end{axis}% 
\end{tikzpicture}
\end{subfigure} 
\begin{subfigure}[]{0.47\linewidth}
\begin{tikzpicture}[scale=0.58]
    \begin{axis}[xlabel=$x$,
                 ylabel=Pressure,
                 label style={font=\large},
                 legend pos=south west,
                 %legend entries={HF,$k=24$ Div,$k=102$ Div,$k=201$ Div,$k=24$ Skew,$k=102$ Skew,$k=201$ Skew},
                 legend style={font=\large},
                 grid=both,
                 ticks=both,
                 width=1.4\linewidth, 
                 height=1.0\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis]
                 \addplot[color=black,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_exact_1000.txt};
                 \addplot[color=red,style=dashed,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_approx_unstable_24_1000.txt};
                 \addplot[color=blue,style=dashed,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_approx_unstable_102_1000.txt};
                 \addplot[color=black!50!green,style=dashed,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_approx_unstable_201_1000.txt};
                 \addplot[color=red,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_approx_24_1000.txt};
                 \addplot[color=blue,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_approx_102_1000.txt};
                 \addplot[color=black!50!green,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_approx_201_1000.txt};
    \end{axis}% 
\end{tikzpicture}
\end{subfigure}

\begin{subfigure}[]{0.47\linewidth}
\begin{tikzpicture}[scale=0.58]
    \begin{axis}[xlabel=$x$,
                 ylabel=Density,
                 label style={font=\large},
                 legend pos=south west,
                 legend entries={HF,$k=24$ Div,$k=102$ Div,$k=201$ Div,$k=24$ Skew,$k=102$ Skew,$k=201$ Skew},
                 legend style={font=\large},
                 grid=both,
                 ticks=both,
                 width=1.4\linewidth, 
                 height=1.0\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis]
                 \addplot[color=black,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_exact_3000.txt};
                 \addplot[color=red,style=dashed,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_approx_unstable_24_3000.txt};
                 \addplot[color=blue,style=dashed,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_approx_unstable_102_3000.txt};
                 \addplot[color=black!50!green,style=dashed,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_approx_unstable_201_3000.txt};
                 \addplot[color=red,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_approx_24_3000.txt};
                 \addplot[color=blue,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_approx_102_3000.txt};
                 \addplot[color=black!50!green,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/density_approx_201_3000.txt};
    \end{axis}% 
\end{tikzpicture}
\caption{}
\label{density_reduction_1D}
\end{subfigure} 
\begin{subfigure}[]{0.47\linewidth}
\begin{tikzpicture}[scale=0.58]
    \begin{axis}[xlabel=$x$,
                 ylabel=Pressure,
                 label style={font=\large},
                 legend pos=south west,
                 %legend entries={HF,$k=24$ Div,$k=102$ Div,$k=201$ Div,$k=24$ Skew,$k=102$ Skew,$k=201$ Skew},
                 legend style={font=\large},
                 grid=both,
                 ticks=both,
                 width=1.4\linewidth, 
                 height=1.0\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis]
                 \addplot[color=black,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_exact_3000.txt};
                 \addplot[color=red,style=dashed,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_approx_unstable_24_3000.txt};
                 \addplot[color=blue,style=dashed,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_approx_unstable_102_3000.txt};
                 \addplot[color=black!50!green,style=dashed,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_approx_unstable_201_3000.txt};
                 \addplot[color=red,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_approx_24_3000.txt};
                 \addplot[color=blue,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_approx_102_3000.txt};
                 \addplot[color=black!50!green,style=solid,style=ultra thick]  table[x = time, y = energy] {./data/Compressible_Euler/1D/Snapshots/pressure_approx_201_3000.txt};
    \end{axis}% 
\end{tikzpicture}
\caption{}
\label{pressure_reduction_1D}
\end{subfigure}
\caption{Qualitative comparison between different formulations for the reduced model in terms of density (\protect\subref{density_reduction_1D})  and pressure(\protect\subref{pressure_reduction_1D}) at $t=0.1,0.3$ and $1$s. Results for the advective formulation are not showed here because the related reduced solutions are unstable after few time steps.}
\end{figure}

\begin{figure}[h!]
\centering
\begin{subfigure}[]{0.47\linewidth}
\begin{tikzpicture}[scale=0.58]
    \begin{axis}[ylabel = e(T),
                 xlabel=Basis DEIM,
                 label style={font=\large},
                 legend pos=north east,
                 legend entries={POD+DEIM,POD},
                 legend style={font=\large},
                 grid=both,
                 ticks=both,
                 width=1.4\linewidth, 
                 height=1\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
%                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis]
                 \addplot[color=blue,style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/error_DEIM/error_DEIM.txt};      
                 \addplot[color=red,style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/error_DEIM/error_POD.txt};
    \end{axis}
\end{tikzpicture}
\caption{}
\label{error_DEIM}
\end{subfigure}
\begin{subfigure}[]{0.47\linewidth}
\begin{tikzpicture}[scale=0.58]
    \begin{axis}[ylabel = Energy(T),
                 xlabel=Basis DEIM,
                 label style={font=\large},
                 legend pos=north east,
                 legend entries={POD+DEIM,POD},
                 legend style={font=\large},
                 grid=both,
                 ticks=both,
                 width=1.4\linewidth, 
                 height=1\linewidth,
                 minor x tick num=1,
                 minor y tick num=2,	
%                 yticklabel style={/pgf/number format/.cd,fixed,precision=9},
                 scaled x ticks = true,
                 enlargelimits=false,
                 scale only axis]
                 \addplot[color=blue,style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/error_DEIM/error_energy_DEIM.txt};      
                 \addplot[color=red,style=solid,style=ultra thick]  table[x = time, y = quantity] {./data/Compressible_Euler/1D/error_DEIM/error_energy_POD.txt};
    \end{axis}
\end{tikzpicture}
\caption{}
\label{energy_DEIM}
\end{subfigure}
\caption{Comparison between standard POD and POD with DEIM treatment of the nonlinear term in terms of the error (\protect\subref{error_DEIM}) and the total energy (\protect\subref{energy_DEIM}).}
\label{fig:RIC2}
\end{figure}

\subsection{Continuous Variable Resonance Combustor}

CVRC is a model rocket combustor designed and operated at Purdue University (Indiana, U.S.) to investigate combustion instabilities \cite{yu2008combustion}. This setup is called the Continuously Variable Resonance Combustor (CVRC) because the length of the oxidizer injector can be varied continuously, allowing for a detailed investigation of the coupling between acoustics and combustion in the chamber \cite{garby2013simulations}. The 2D/3D high-fidelity simulations of CVRC are expensive. Thus to get a fast analysis tool, a quasi-1D model has been proposed by Smith et al. \cite{smith2008computational} and further developed by Frezzotti et al. \cite{frezzotti2015determination,frezzotti2017numerical,frezzotti2018quasi}.

The CVRC consists of three parts: oxidizer post, combustion chamber and exit nozzle, as shown in Fig. \ref{fig:radius}. The oxidizer is injected from the left end of the oxidizer post and meets the fuel that is injected through an annular ring around the oxidizer injector, at the back-step. The combustion happens in a region around the back-step. The combustion products flow through the chamber and exit the system from the nozzle.  Both the injector and the nozzle are operated at choked condition during the experiment. The length of the oxidizer post $L_{op}$ of the CVRC can be varied continuously, leading to different behavior of the combustion stability. In this paper, we will focus on the case with $L_{op}= 14.0$ cm, in which the combustion is unstable.

The geometry parameters of the quasi-1D CVRC with a oxidizer post length  $L_{op}= 14.0$ cm are shown in Table \ref{tab:geometry_parameters}.  The back-step and the converging part of the nozzle are sinusoidally contoured to avoid discontinuity of the radius that will invalidate the quasi-1D governing equations presented in the next subsection.

\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{pic/radius}
	\caption{Geometry of quasi-1D CVRC model.}
	\label{fig:radius}
\end{figure}

\begin{table} [h]
	\centering
	\caption{Geometry parameters of the quasi-1D CVRC with an oxidizer post length $L_{op}=14$ cm.}
	\centering
	\begin{tabular}{c c c c c c }
		\toprule
		\centering
		\multirow{2}{*}{Section} &
		\multicolumn{2}{c}{Oxidizer post} &
		\multirow{2}{*}{Chamber} &
		\multicolumn{2}{c}{Nozzle} \\
		\cmidrule(lr){2-3} \cmidrule(lr){5-6}
		& injector & back-step & & converging part & diverging part\\
		\midrule
		Length (cm) & 12.99 & 1.01 & 38.1 & 1.27 & 3.4 \\
		Radius (cm) & 1.02  & $1.02 \sim 2.25$ & 2.25 & $2.25 \sim 1.04$ & $1.04 \sim 1.95$ \\
		\bottomrule
		\label{tab:geometry_parameters}
	\end{tabular} 
\end{table}

The fuel is pure gaseous methane. The oxidizer is a mixture of 42\% oxygen and 58\% water (per unit mass). The oxidizer is injected in the oxidizer post at a temperature $T_{ox}=1030$ K so that both water and oxygen are in the gaseous phase. The operating conditions are listed in Table \ref{tab:operating-conditions}.

\begin{table} [h]
	\centering
	\caption{CVRC operating conditions.}
	\centering
	\begin{tabular}{l l l}
		\toprule
		\centering
		Parameter & Unit & Value \\
		\midrule
		Fuel mass flow rate, $\dot{m}_{f}$ & kg/s & 0.027   \\
		Fuel temperature, $T_{f}$ & K & 300   \\
		Oxidizer mass flow rate, $\dot{m}_{ox}$ & kg/s & 0.32   \\
		Oxidizer temperature, $T_{ox}$ & K & 1030   \\
		$O_2$ mass fraction in oxidizer, $Y_{O_2}$ & -- & 42.4\%   \\
		$H_2O$ mass fraction in oxidizer, $Y_{H_2O}$ & -- & 57.6\%   \\
		Mean chamber pressure & MPa & 1.34 \\
		Equivalence ratio, $E_r$ & -- & 0.8 \\
		\bottomrule
		\label{tab:operating-conditions}
	\end{tabular} 
\end{table}

For the combustion, we consider the one-step reaction model
\begin{equation*}\label{eq:combustion}
CH_4 + 2O_2 \rightarrow CO_2 + 2H_2O
\end{equation*}
We assume that the fuel reacts instantaneously to form products, allowing us to neglect intermediate species and finite reaction rates. As the equivalence ratio is less than one, there is oxidizer left after the combustion. Therefore, only two species need to be considered: oxidizer and combustion products.

The governing equations that describe the conservation of mass, momentum, and energy of the quasi-1D CVRC flow, are the quasi-1D unsteady Euler equations for multiple species, expressed in conservative form as
\begin{equation*}\label{eq:quasi-1D-equation}
\frac{\partial \mathbf{u}}{\partial t} + \frac{\partial \mathbf{f}}{\partial x} = \mathbf{s}_A + \mathbf{s}_f + \mathbf{s}_q.
\end{equation*}
The conserved variable vector $\mathbf{u}$ and the convective flux vector $\mathbf{f}$ are
\begin{equation*}\label{eq:CVRC-u-f}
\mathbf{u}= \left( \begin{gathered}
\rho A  \\
\rho uA  \\
\rho EA  \\
\rho Y_{ox} A \\
\end{gathered} \right), 
\mathbf{f} = \left( \begin{gathered}
\rho uA  \\
\left(\rho u^2 + p\right)A  \\
\left(\rho E + p\right)uA  \\
\rho uY_{ox} A \\
\end{gathered} \right),
\end{equation*}
where $\rho$ is the density, $u$ is the velocity, $p$ is the pressure, $E$ is the total energy, $Y_{ox}$ is the mass fraction of oxidizer, and $A=A(x)$ is the cross section area of the duct. The pressure $p$ can be computed using the conserved variables as
\begin{equation*}\label{eq:total-engery}
E = \frac{p}{\rho (\gamma - 1)} + \frac{u^2}{2} - C_p T_{ref},
\end{equation*}
where $T_{ref}$ is the reference temperature and is set as 298.15 K in this paper. The temperature $T$ is recovered from the equation of state $p = \rho R T$. The gas properties  $C_p$, $R$ and $\gamma$ are computed as $C_p= \sum C_{pi}Y_i$, $R=\sum R_iY_i$ and $ \gamma= C_p/(C_p-R)$, respectively.

The source terms are
\begin{equation*}\label{eq:source-terms}
\mathbf{s}_A = \left( \begin{gathered}
0  \\
p \frac{dA}{dx}  \\
0  \\
0 \\
\end{gathered} \right), 
\mathbf{s}_f = \left( \begin{gathered}
{\dot \omega}_f  \\
{\dot \omega}_f u  \\
{\dot \omega}_f \left(h_{0}^{f} + \Delta h_{0}^{rel} \right)  \\
{\dot \omega}_{ox} \\
\end{gathered} \right), 
{\mathbf{s}_q} = \left( \begin{gathered}
0  \\
0  \\
q'  \\
0 \\
\end{gathered} \right),
\end{equation*}
where $\dot{\omega}_f$ is the depletion rate of the fuel, $\dot{\omega}_{ox}$ is the depletion rate of the oxidizer, $h_0^f$ is the total enthalpy of the fuel, $\Delta h_{0}^{rel}$ is the heat of reaction per unit mass of fuel and $q'$ is the unsteady heat release term. $\mathbf{s}_A$ accounts for area variations, $\mathbf{s}_f$ and $\mathbf{s}_q$ are related to the combustion. $\mathbf{s}_f$ represents the addition of the fuel and its combustion with the oxidizer, which in turn results in the creation of the combustion products. The depletion rate of the fuel is
\begin{equation}\label{eq:wf}
\dot{\omega}_{f}= \frac{k_f \dot{m}_f Y_{ox} \left(1+sin\xi\right)}{l_f-l_s},
\end{equation} 
where
\begin{equation*}\label{eq:xi}
\xi= -\frac{\pi}{2} + 2\pi\frac{x-l_s}{l_f-l_s}, \hspace{0.5cm} \forall \hspace{0.2cm} l_s < x < l_f.
\end{equation*}
The setting of the fuel injection restricts the combustion to the region $l_s < x < l_f$. The reaction constant $k_f$ is selected to insure that the fuel is consumed within the specified combustion zone. The depletion rate of the oxidizer is computed by 
\begin{equation*}\label{eq:wox}
\dot{\omega}_{ox} = C_{o/f} \dot{\omega}_f,
\end{equation*}
where $C_{o/f}$ is the oxidizer-to-fuel ratio.

The unsteady heat release term $q'$, also called the combustion response function, models the coupling between acoustics and combustion. In this paper, we use the combustion response function designed by Frezzotti et al. \cite{frezzotti2017numerical,frezzotti2018quasi}, which is a function of the velocity, sampled at specific abscissa $\hat{x}$ that is almost coincident with the antinode of the first longitudinal modal shape, with a certain time lag $\tau$, i.e.,
\begin{equation}\label{eq:response-function}
q'\left( x,t\right) = \alpha g\left(x\right)  A\left(x\right) \left[ u\left( \hat{x},t-\tau \right) - \bar{u}\left( \hat{x} \right) \right].
\end{equation}
Here $\bar{u}$ is the time averaged velocity, estimated with the steady-state quasi-1D model assuming $q'=0$, and $g(x)$ is a Gaussian distribution  
\begin{equation*}\label{eq:gx}
g\left(x\right)= \frac{e^{-\frac{\left(x-\mu\right)^2}{2\sigma^2}}}{\sqrt{2\pi\sigma^2}},
\end{equation*}
where $\mu$ is the mean and $\sigma$ is the standard deviation. The amount of heat release due to velocity oscillations is controlled by the parameter $\alpha$.

The boundary conditions for the quasi-1D CVRC flow include the fixed mass flow rate and the stagnation temperature at the head-end of the oxidizer injector, and the supersonic outflow at the exit of the nozzle.

Prior to unsteady simulation, the quasi-1D CVRC needs to be excited, which can be achieved by adding a perturbation to the steady-state solution. The perturbation is added by forcing the mass flow rate with a multi-sine signal
\begin{equation}\label{eq:multisine}
\dot{m}_{ox} \left(t\right)= \dot{m}_{ox,0} \left[1 + \delta\sum_{k=1}^{K}  sin\left(2\pi k\Delta f t\right) \right],
\end{equation}
where $\dot{m}_{ox,0}$ is the oxidizer mass flow rate in Table \ref{tab:operating-conditions}, $\Delta f$ is the frequency resolution and $K$ is the number of frequencies. In this paper, $\Delta f = 50 $ Hz and $K=140$, resulting in a minimal frequency of 50 Hz and a maximal frequency of 7000 Hz. $\delta$ is required to be small to control the amplitude of the perturbation and is set as 0.1\%.

The procedure of the unsteady simulation of the quasi-1D CVRC flow includes three steps:
\begin{enumerate}[(1)]
	\item  Compute the steady-state solution by setting $\dot{m}_{ox}=\dot{m}_{ox,0} $ and $q'=0$.
	\item  Excite the system by adding a perturbation to the oxidizer mass flow rate according to (\ref{eq:multisine}) and setting $q'=0$.
	\item  Perform the unsteady simulation by turning on the combustion response function $q'$ in (\ref{eq:response-function}) and turning off the oxidizer mass flow rate perturbation by setting $\dot{m}_{ox}=\dot{m}_{ox,0} $.	
\end{enumerate}

A high-fidelity quasi-1D CVRC flow solver is built by employing the MUSCL reconstruction \cite{delis1998tvd}, the Lax-Friedrichs flux \cite{shu1988efficient} and the strong stability preserving, three-stage Runge-Kutta (SSP RK3) time stepping \cite{jiang1996efficient}.