### 2018-07-12 paper.tex
first section of results corrected

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a contributed volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[graybox]{svmult}

% choose options for [] as required from the list
% in the Reference Guide

\usepackage{mathptmx}       % selects Times Roman as basic font
\usepackage{helvet}         % selects Helvetica as sans-serif font
\usepackage{courier}        % selects Courier as typewriter font
\usepackage{type1cm}        % activate if the above 3 fonts are
                            % not available on your system
%
\usepackage{makeidx}         % allows index generation
\usepackage{graphicx}        % standard LaTeX graphics tool
                             % when including figure files
\usepackage{epstopdf}

\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{subcaption}
\captionsetup{compatibility=false} % otherwise it raises some issues regarding compatibility

\usepackage{multicol}        % used for the two-column index
\usepackage[bottom]{footmisc}% places footnotes at page bottom
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
%\usepackage[linesnumbered, ruled, vlined]{algorithm2e}
%\usepackage{algorithm}% http://ctan.org/pkg/algorithms
%\usepackage{algorithmicx} 
%\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
%\renewcommand{\algorithmicensure}{\textbf{Output:}}
\usepackage{enumerate}
\usepackage{multirow}
\usepackage{cleveref}

\newcommand{\commnet}{\color{red}}

\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\pgfplotsset{
    % most recent feature set of pgfplots
    compat=newest,
    % some settings for grid
    grid style={black!20!white, thin, densely dotted},
    % modify plot appearance
    every axis plot/.append style={no markers, thick},
    % i like the labels a bit smaller
    label style={font=\small},
    tick label style={font=\small}
}
\usepgfplotslibrary{external} 
\tikzexternalize[prefix=TikzFigures/]

%\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{float} % Required for tables and figures in the multi-column environment - they need to be placed in specific locations with the [H] (e.g. \begin{table}[H])
%\usepackage{hyperref} % For hyperlinks in the PDF

% see the list of further useful packages
% in the Reference Guide

\makeindex             % used for the subject index
                       % please use the style svind.ist with
                       % your makeindex program

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title*{Conservative Model Order Reduction of Fluid Flow}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{Bababk Maboudi Afkham, Nicol\`o Ripamonti, Qian Wang, Jan S. Hesthaven and Karen E. Willcox}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Name of First Author \at Name, Address of Institute, \email{name@email.address}
\and Name of Second Author \at Name, Address of Institute \email{name@email.address}}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

%\abstract*{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content. The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access. This allows unregistered users to read the abstract as a teaser for the complete chapter. As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.
%Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf. source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript. Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}

%\abstract{In the context of model order reduction, conservation of stability for systems of hyperbolic partial differential equations, remains a challenge. Recently preserving structures, invariants and conservation laws, in order to help with stability, and ensure robustness in long-time integrations, has been an area of active research. Energy conservation is a fundamental feature of fluid flows which is often violated in the course of model reduction. In this paper we discuss a model reduction routine, that exploits skew-symmetry of conservative and centered discretization schemes, to recover conservation of energy at the level of reduced system. This results in an, overall, correct evolution of the fluid and ensures robustness of the reduced system. We evaluate the performance of the proposed method through numerical simulation of various fluid flows, and also in a numerical simulation of continuous variable resonance combustor model.}

\abstract{In the past decade, model order reduction (MOR) has been successful in reducing the computational complexity of elliptic and parabolic systems of partial differential equations (PDEs). However, MOR of hyperbolic equations remains a challenge. Symmetries and conservation laws, which are a distinctive feature of such systems, are often destroyed over the course of conventional MOR techniques. This often result in a perturbed, and often unstable reduced system. The importance of the conservation of energy is well-known in a correct numerical integration of fluid flow. In this paper, we discuss a model reduction routine, that exploits skew-symmetry of conservative and centered discretization schemes, to recover conservation of energy at the level of reduced system. Moreover, we argue that the reduced system, constructed with the new method, can be identified by a reduced energy that mimics the energy of the high-fidelity system. Therefore, the loss in energy, associated with model reduction, remains constant in time. This results in an, overall, correct evolution of the fluid and ensures robustness of the reduced system. We evaluate the performance of the proposed method through numerical simulation of various fluid flows, and also in a numerical simulation of continuous variable resonance combustor model.}

% sections are added here

\input{./sections/1.intro.tex}
\input{./sections/2.mor.tex}
\input{./sections/3.skew.tex}
\input{./sections/4.mor_skew.tex}
\input{./sections/5.res.tex}
\input{./sections/6.con.tex}

\bibliographystyle{plain}
\bibliography{references}

\end{document}