### 2018-07-12 sections/1.intro.tex

\section{Introduction} \label{sec:intro}

Model order reduction (MOR), and in particular reduced basis (RB) methods, has emerged as a powerful approach to cope with the complex and computationally intensive models in engineering and science. Such techniques, construct a reduced ordered representation for the state of a model which accurately approximates the configuration of the system. The evaluation of the this representation is, then, possible with considerable acceleration.

Although RB methods are successful in reducing the computational complexity of models with elliptic and parabolic partial differential equations (PDEs), MOR of systems of hyperbolic equations, or models with strong advective terms, remains a challenge. Often, such models arise from a set of invariants and conservation laws. Such symmetries are violated over the course of MOR which result in a qualitatively wrong, and sometimes unstable, solution.

Constructing MOR techniques and RB methods that preserve intrinsic structures has recently attracted attention \cite{doi:10.1137/17M1111991,1705.00498,kalashnikova2014stabilization,farhat2015structure,doi:10.1137/110836742,beattie2011structure,doi:10.1137/140978922}. Structure preservation can recover a physically meaningful reduced model, rather than a pure algebraic coupling of equations. This enforces robustness and can help with the stability of reduced model. A non-exhaustive list of recent works on structure preserving MOR is as follows. Preserving time-symmetries of Lagrangian, Hamiltonian, and port-Hamiltonian systems can be found in the works of \cite{Carlberg:2014ky,doi:10.1137/140978922,doi:10.1137/17M1111991,1705.00498,beattie2011structure,chaturantabut2016structure,gugercin2012structure}. Conserving the inf-sup stability, in the context of finite element methods, can be found in the works of \cite{farhat2015structure,ballarin2015supremizer}. Furthermore, a flux preserving model reduction for finite volume methods is presented in \cite{carlberg2018conservative}.

Large scale simulation of fluid flows appear in a wide range of disciplines and industries. Therefore, MOR of fluid flows, specially when advective terms are dominant, is of a high importance. It is well known that conservation of the energy, specially kinetic energy, is essential for a qualitatively correct numerical integration of fluid flows. Conventional model reduction techniques often violates conservation of mass, momentum \cite{carlberg2018conservative}, or energy in fluid flows which result in an unstable reduced system, specially over long time-integration. The method discussed in \cite{carlberg2018conservative} conserves the mass and momentum for a finite volume discretization scheme. However, a method that also conserves the energy of the fluid flow is not known to the authors.

Skew-symmetric forms for fluid flows are centered time and space discretization schemes, typically using a finite differences discretization method, that recovers time-symmetries of a fluid at the discrete level. Such discretization schemes are studies comprehensively over the past few decades and can be found in the works of \cite{morinishi2010skew,morinishi1998fully,desjardins2008high,reiss2014conservative,reiss2014conservative} and the references therein. These schemes construct a skew-symmetric differential operator, acting on the momentum vector field, that ensures conservation quadratic invariants, such as energy.

In this paper we discuss how to preserve skew-symmetry of differential operators at the level of reduced system. This results in conservation of quadratic invariants. Furthermore, we show that the reduced system, as a system of coupled differential equations, contains quadratic invariants and an associated energy which approximates the energy of the high-fidelity system. Therefore, a proper time stepping scheme can preserve the reduced representation of the energy, and therefore, the loss in energy due to model reduction remains constant in time.

The rest of this paper is organized as follows. In \Cref{sec:mor} we summarize the theory on MOR and introduce proper orthogonal decomposition (POD) as a conventional RB method. We discuss skew-symmetric and conservatives methods for compressible and incompressible fluid flows in \Cref{sec:skew}. Conservative and energy-preserving model reduction of fluid flows is discussed in \Cref{sec:mor_skew}. We evaluate the performance of the method, in numerical simulations of incompressible and compressible fluid flow in \Cref{sec:res}. We also apply the method to construct a reduced system for the continuous variable resonance combustor, a one dimensional reaction-diffusion model for a rocket engine. Finally, we present conclusive remarks in \Cref{sec:con}.